---
title: "Case Study 1: Bike Trip Data"
author: "Rani Bhangu"
date: '2023-03-18'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Installing the package Tidyverse
We require the package Tidyverse for any sort of Data operations in R. Installing and loading the library using the following:

```{r Tidyverse,echo=TRUE}
install.packages("tidyverse")
library(tidyverse)
```

# Reading the Data
Uploaded the Data file(CSV file) in R environment and then stored it into a dataframe called df by using the read_csv function

```{r read_csv, echo=TRUE}
df=read_csv("202204-divvy-tripdata.csv")
```

# Summary of the Data
Initially we observed all the details about the columns. The total number of rows also was obtained from the summary.

```{r summary}
summary(df)
```

# Segregate the member data
To seperate the annual members we create a new dataframe called df2 which will have only the member values from member_casual column and also we find the number of rows of df2

```{r members}
df2=df[df$member_casual=='member',]
nrow(df2)
```

# Calculating the time taken for Bike rides
We create using the mutate function a new dataframe df3 containing the time difference column in hours. The time difference is taken by the difftime function using start time and end time of the rides

```{r difftime}
df3=mutate(df,time_difference_hours = difftime(ended_at, started_at, units = "hours"))
```

# Analying the time differences obtained
We observed the time differences for all the rides and found the mean time taken for a ride and also the range i.e min and max values.

```{r time_difference}
df3$time_difference_hours
range(df3$time_difference_hours)
mean(df3$time_difference_hours)
```

# Installing Geosphere
To calculate the distance between the start station and end station we use the disthaversine function which is a part of geosphere library. So installing the geosphere library.

```{r}
install.packages("geosphere")
library(geosphere)
```

# Creating a dataframe for the distances
Using disthaversine function we calculate the distance and we store it in the new dataframe called df4

```{r distHaversine}
df4=mutate(df,distance_km = distHaversine(cbind(start_lng, start_lat), cbind(end_lng, end_lat))*0.001)
```

# Analyzing the ride distances
We observe the distance_km data as well as summarize it

```{r ride_distances}
df4$distance_km
summary(df4$distance_km)
```

